Plug-and-play dual-tree algorithm runtime analysis

### 2014-12-01 bounds-files/sec01-supporting-lemmas.tex
clarified proof of lemma 1

\begin{lemma}
Consider a dataset $S$ with expansion constant $c$ and a subset $C \subseteq S$
such that every point in $C$ is separated by $\delta$.  Then, given any query
point $p \not\in S$ and some radius $\rho \delta$,

\begin{equation}
| B_S(p, \rho \delta) \cap C | \le c^{2 + \log_2 \rho}.
\end{equation}
\label{lem:packing}
\end{lemma}

\begin{proof}
This is based on the packing argument from Lemma 4.1 in \cite{langford2006}.
Observe that $B_S(p, \rho \delta) \subseteq B_S(p_i, 2 \rho \delta)$ for any
$p_i \in S$, and that $| B_S(p, 2 \rho \delta) | = c^{2 + \log_2 \rho} |
B_S(p, \delta / 2) |$.  Because each point in $C$ is separated by $\delta$, the
number of points in $B_S(p, \rho \delta) \cap C$ is
bounded by the number of disjoint balls of radius $\delta / 2$ that can be
packed into $B_S(p, \rho \delta)$.  In the worst case, this packing is
perfect, and

\begin{equation}
|B_S(p, \rho \delta)| \le \frac{|B_S(p_i, 2 \rho \delta)|}{|B_S(p_i, \delta
/ 2)|} \le c^{2 + \log_2 \rho}.
\end{equation}
\end{proof}

>>>>

\begin{lemma}
Consider a dataset $S$ with expansion constant $c$ and a subset $C \subseteq S$
such that every point in $C$ is separated by $\delta$.  Then, for any point 
$p \not\in S$ and any radius $\rho \delta > 0$:
\begin{equation}
| B_S(p, \rho \delta) \cap C | \le c^{2 + \lceil \log_2 \rho \rceil}.
\end{equation}
\label{lem:packing}
\end{lemma}

\begin{proof}
The proof is based on the packing argument from Lemma 4.1 in 
\cite{langford2006}. Consider two cases: first, let $d(p, p_i) > \rho \delta$ 
for any $p_i \in S$. In this case, $B_S(p, \rho \delta) = \emptyset$ and the 
lemma holds trivially. 
Otherwise, let $p_i \in S$ be a point such that $d(p, p_i) \leq \rho \delta$.
Observe that $B_S(p, \rho \delta) \subseteq B_S(p_i, 2 \rho \delta)$. 
Also, $| B_S(p_i, 2 \rho \delta) | = c^{2 + \lceil \log_2 \rho 
\rceil} | B_S(p_i, \delta / 2) |$ by the definition of the expansion constant.
Because each point in $C$ is separated by $\delta$, the
number of points in $B_S(p, \rho \delta) \cap C$ is
bounded by the number of disjoint balls of radius $\delta / 2$ that can be
packed into $B_S(p, \rho \delta)$.  In the worst case, this packing is
perfect, and

\begin{equation}
|B_S(p, \rho \delta)| \le \frac{|B_S(p_i, 2 \rho \delta)|}{|B_S(p_i, \delta
/ 2)|} \le c^{2 + \lceil \log_2 \rho \rceil}.
\end{equation}
\end{proof}
